\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-07-18 paper.qmd
Update Digit 1 images

<!-- Fit the model and compute error-->
```{r}
## Compute hexbin parameters
num_bins_x_mnist <- 37
lim1 <- mnist_scaled_obj$lim1
lim2 <- mnist_scaled_obj$lim2
r2_mnist <- diff(lim2)/diff(lim1)

@@

<!--bin1 = 37, bin2 = 26, b = 962, non_empty = 398-->
```{r}
#| echo: false
#| fig-cap: "(a) Model generated in $2\\text{-}D$, and (b) $p\\text{-}D$ model error in $2\\text{-}D$. The $2\\text{-}D$ model shows a non-linear continuous structure. Most low $p\\text{-}D$ model errors are distributed along the lower edge of the $2\\text{-}D$ structure, while most high p-D model errors are concentrated along the upper edge."
#| label: fig-model-mnist
#| fig-pos: H

@@

There are certain data points that exhibit high error rates due to their deviation from the usual \pD{} data structure, which makes them anomalies (@fig-model-mnist (b)). These anomalies can be classified into two types: those that are anomalies within the non-linear structure and those that lie outside of it. The images associated with high model error points within the non-linear structure display different patterns of the digit 1, as shown in @fig-mnist-anomalies (a). However, when comparing these images to the ones found outside of the non-linear structure, it becomes evident that the latter display different patterns of the digit 1 (@fig-mnist-anomalies (b)).

<!--images that occur large error-->
<!-- within the nonlinear structure-->
```{r}
## Data with pixel values
mnist_data <- read_rds("data/mnist/mnist_digit_1.rds")

pixels_gathered <-  mnist_data |>
  mutate(instance = row_number()) |>
  gather(pixel, value, -Label, -instance) |>
  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
  filter(instance %in% c( 4699, 4785, 5173, 6671, 2186, 4636, 4882, 
                          773, 2825, 6169, 6948, 5987, 1282, 7715, 
                          622, 2989, 6912, 4370, 1227, 4042, 2444, 
                          836, 1654))

imge_error_sample_within <- pixels_gathered |> 
  ggplot(aes(x, y, fill = value)) +  
  geom_tile() +  
  facet_wrap(~ factor(instance, levels = c( 4699, 4785, 5173, 6671, 2186, 4636, 
                                            4882, 773, 2825, 6169, 6948, 5987, 
                                            1282, 7715, 622, 2989, 6912, 4370, 
                                            1227, 4042, 2444, 836, 1654)), 
             ncol = 4) +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none")
```

<!--outside the nonlinear structure-->
```{r}
img_error <- error_df |> 
  filter(row_wise_abs_error >= 8) |> 
  filter((!ID %in% c(4699, 4785, 5173, 6671, 2186, 4636, 4882, 773, 
                            2825, 6169, 6948, 5987, 1282, 7715, 622, 2989, 
                            6912, 4370, 1227, 4042, 2444, 836, 1654))) |>
  pull(ID)

pixels_gathered <-  mnist_data |>
  mutate(instance = row_number()) |>
  gather(pixel, value, -Label, -instance) |>
  extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
  filter(instance %in% img_error)

imge_error_sample <- pixels_gathered |> 
  ggplot(aes(x, y, fill = value)) +  
  geom_tile() +  
  facet_wrap(~ instance, ncol = 2) +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none") 
```

```{r}
#| echo: false
#| fig-cap: "Some images of handwritten digit 1 which occur high model error (a) within the non-linear strcuture, and (b) outside the non-linear structure. The images shows different patterns of digit 1."
#| label: fig-mnist-anomalies
#| out-height: 35%
#| fig-pos: H

imge_error_sample_within + imge_error_sample +
  plot_layout(ncol=2) +
  plot_annotation(tag_levels = 'a')
```

>>>>

<!-- Fit the model and compute error-->
```{r}
## Compute hexbin parameters
num_bins_x_mnist <- 22
lim1 <- mnist_scaled_obj$lim1
lim2 <- mnist_scaled_obj$lim2
r2_mnist <- diff(lim2)/diff(lim1)

@@

<!--bin1 = 22, bin2 = 17, b = 374, non_empty = 140-->
```{r}
#| echo: false
#| fig-cap: "(a) Model generated in $2\\text{-}D$, and (b) $p\\text{-}D$ model error in $2\\text{-}D$. The $2\\text{-}D$ model shows a non-linear continuous structure. Most low $p\\text{-}D$ model errors are distributed along the lower edge of the $2\\text{-}D$ structure, while most high p-D model errors are concentrated along the upper edge."
#| label: fig-model-mnist
#| fig-pos: H

@@

There are certain data points that exhibit high error rates due to their deviation from the usual \pD{} data structure, which makes them anomalies (@fig-model-mnist b). These anomalies can be classified into two types: those that are anomalies within the non-linear structure and those that lie outside of it. The images associated with high model error points within the non-linear structure display different patterns of the digit 1, as shown in @fig-mnist1-within. However, when comparing these images to the ones found outside of the non-linear structure, it becomes evident that the latter display different patterns of the digit 1 (@fig-mnist1-out).

<!--images that occur large error-->
<!-- within the nonlinear structure-->
<!--outside the nonlinear structure-->

::: {#fig-mnist-anomalies layout-ncol="2" fig-pos="H"}
![](figures/mnist/img_error_sample_within.png){#fig-mnist1-within}

![](figures/mnist/img_error_sample_out.png){#fig-mnist1-out}

Some images of handwritten digit 1 which occur high model error (a) within the non-linear structure, and (b) outside the non-linear structure. The images shows different patterns of digit 1.
:::

### 2024-07-18 paper.tex
Update Digit 1 images

There are certain data points that exhibit high error rates due to their
deviation from the usual \pD{} data structure, which makes them
anomalies (Figure~\ref{fig-model-mnist} (b)). These anomalies can be
classified into two types: those that are anomalies within the
non-linear structure and those that lie outside of it. The images
associated with high model error points within the non-linear structure
display different patterns of the digit 1, as shown in
Figure~\ref{fig-mnist-anomalies} (a). However, when comparing these
images to the ones found outside of the non-linear structure, it becomes
evident that the latter display different patterns of the digit 1
(Figure~\ref{fig-mnist-anomalies} (b)).

@@

@@

\includegraphics[width=1\textwidth,height=0.35\textheight]{paper_files/figure-pdf/fig-mnist-anomalies-1.pdf}

@@

\caption{\label{fig-mnist-anomalies}Some images of handwritten digit 1
which occur high model error (a) within the non-linear strcuture, and
(b) outside the non-linear structure. The images shows different
patterns of digit 1.}

>>>>

There are certain data points that exhibit high error rates due to their
deviation from the usual \pD{} data structure, which makes them
anomalies (Figure~\ref{fig-model-mnist} b). These anomalies can be
classified into two types: those that are anomalies within the
non-linear structure and those that lie outside of it. The images
associated with high model error points within the non-linear structure
display different patterns of the digit 1, as shown in
Figure~\ref{fig-mnist1-within}. However, when comparing these images to
the ones found outside of the non-linear structure, it becomes evident
that the latter display different patterns of the digit 1
(Figure~\ref{fig-mnist1-out}).

@@

\begin{minipage}{0.50\linewidth}

@@

\includegraphics{figures/mnist/img_error_sample_within.png}

@@

\subcaption{\label{fig-mnist1-within}}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\includegraphics{figures/mnist/img_error_sample_out.png}

\subcaption{\label{fig-mnist1-out}}

\caption{\label{fig-mnist-anomalies}Some images of handwritten digit 1
which occur high model error (a) within the non-linear structure, and
(b) outside the non-linear structure. The images shows different
patterns of digit 1.}
